---
title: "bueno-gabrielle-ada-homework-1"
output:
  html_document:
    toc: yes
    toc_float: true
    toc_depth: 2
    df_print: paged
---

# Challenge 1

1. Load file
```{r}
f <- "/Users/gabriellebueno/RStudio/Repositories/ada-2021-datasets/darwin.txt"
darwin <- read.delim(f, header = TRUE, sep = "\t")
```

2. Calculate and print out how many paragraphs are there in the chapter (do not count the chapter title as a paragraph)
```{r}
lengths(darwin)
#A: 56
```

3. Print out the 34th paragraph
```{r}
p <- darwin[34, ]
p
```

4. Split this vector (note: I'm assuming you mean the entire darwin text, and not just the 34th paragraph--the wording of the assignment is a bit unclear)
```{r}
library(stringr)
DarwinPunct <- gsub("[[:punct:]]","",darwin) ## Removing punctuation
SplitOriginal <- str_split(DarwinPunct, " ") 
split <- SplitOriginal[[1]] #Subsetting the wanted character vector from the list SplitOriginal 
```

5a. How many unique words are their in this vector? 

```{r}
checkUnique <- unique(split)
length(checkUnique)
##A: 2081
```

5b. What is the most common word?  
```{r}
library(DescTools)
Mode(split)
#A: "the", appears 718 times
```

5c. How many words appear 1 time?

```{r}
frequency <- data.frame(table(split))
subsetOne <- frequency[frequency$Freq == 1, ]
lengths(subsetOne)
#A: 1088
```

5d. How many words appear 5 or more times?
```{r}
subsetTwo <- frequency[frequency$Freq >= 5, ]
lengths(subsetTwo)
#A: 369
```

6. Extract the entire text of the last paragraph to a new variable, final_quote:
```{r}
final_quote <- darwin[56, ]
```

7. Split this vector into a new vector, split_quote, comprising a vector of 1-word character strings

```{r}
library(stringr)
#Not removing punctuation here, it shouldn't impact alphabetizing
splitQuoteOriginal <- str_split(final_quote, " ") 
split_quote <- splitQuoteOriginal[[1]] #Subsetting the wanted character vector from the list SplitOriginal 
```

8. Extract every third element from this vector, assign them to a new vector, every_third, and print out the new vector.

```{r}
every_third <- split_quote[seq(1, length(split_quote), 3)]
every_third
```

9. Rearrange the elements of the vector every_third into reverse alphabetical order and print it out again.

```{r}
RevAlph <- sort(every_third, decreasing = TRUE)
RevAlph
```

# Challenge 2

1. Use the function `c()` to create a vector with the average high temperatures in January for Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto, which are 35, 88, 42, 84, 81, and 30 degrees Fahrenheit. Call the object t.
```{r}
t <- c(35, 88, 42, 84, 81, 30)
```

2. Now create a vector with the city names and call the object city.
```{r}
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
```

3. Use the `names()` function and the objects defined in the previous exercises to associate the temperature data with its corresponding city.
```{r}
names(t) <- city
```


4. Use the `[` and `:` operators to access the temperature of the first three cities on the list.
```{r}
t[1:3]
```

5. Use the `[` operator to access the temperatures of Paris and San Juan.
```{r}
t["Paris"]
t["San Juan"]
```

# Challenge 3

1. Create a 8 row by 20 column matrix, named m1, consisting of integers arranged in descending order from 159 to 1. Fill your matrix column-wise.
```{r}
m1 <- matrix(
  data = c(sort(1:159, decreasing = TRUE)),
  nrow = 8,
  ncol = 20
)
m1
```

2. Using single bracket notation, extract and print the element in row 5, column 2.
```{r}
m1[5,2]
#A: 147
```

3. Using single bracket notation, extract and print the contents of rows 5 to 7 of the matrix.
```{r}
m1[5:7, ]
```

4. Using single bracket notation, extract the contents of (row 3, column 4) to (row 6, column 9) as a new variable, m2. What is the class of this variable? What is the mode of this variable?
```{r}
m2 <- m1[3:6, 4:9]
class(m2) #A: "matrix" "array"
mode(m2)
#A: "numeric"
```

# Challenge 4

* (*Technically the second "Challenge 3" on the HW page*)

1. Construct a 4-dimensional, 400-element array (5 x 5 x 4 x 4), named a, consisting of even integers from 2 to 800 (i.e., a descending series)
   + **Question: wouldn't "descending series" imply it should be going down from 800 to 2? I will follow the instructions and do 2 to 800 but wanted to make a note in case that was incorrect.**
```{r}
s <- seq(from = 2, to = 800, by = 2)
a <- array(data = s, dim = c(5, 5, 4, 4))
```

2. Given this matrix, what would the following return?

2a. a[1, 1, 1, 2]
```{r}
a[1, 1, 1, 2] 
```

2b. a[2, 3, 2, ]
```{r}
a[2, 3, 2, ] 
```

2c. a[1:5, 1:5, 3, 3]
```{r}
a[1:5, 1:5, 3, 3]
```

# Challenge 5 

* (*Technically "Challenge 4" on the HW page, but there were two Challenge 3s*)

1. Create a list representing the (simplified) primate taxonomy outlined below and shown in the figure. Name each internal node in the list (i.e., Superfamily and above) with the appropriate taxonomic level and name. Within each superfamily, you can use a vector to store family names (e.g., Superfamily Lorisoidea can be represented as `c("Lorisidae, "Galagidae"))`.
```{r}
# Making the superfamilies into vectors
Lorisoidea <- c("Lorisoidea", "Galagidae")
Lemuroidea <- c("Cheirogaleidae", "Lepilemuridea", "Indriidae", "Lemuridae", "Daubentoniidae")
Tarsioidea <- c("Tarsiidae")
Ceboidea <- c("Cebidae", "Atelidae", "Pitheciidae")
Hominoidea <- c("Hylobatidae", "Hominidae")
Cercopithecoidea <- c("Cercopithecidae")

# Making the parvorders into lists 
Platyrrhini <- list(Ceboidea)
  names(Platyrrhini) <- c("Ceboidea") # I'm sure there's a fancier way to name them, but I had trouble figuring it out, so I'm doing them all one by one.
Catarrhini <- list(Hominoidea, Cercopithecoidea)
  names(Catarrhini) <-c("Hominoidea", "Cercopithecoidea")

# Making the infraorders into lists containing their superfamilies (or parvorders)
Lorisiformes <- list(Lorisoidea)
  names(Lorisiformes) <- c("Lorisoidea") 
Lemuriformes <- list(Lemuroidea)
  names(Lemuriformes) <- c("Lemuroidea")
Tarsiiformes <- list(Tarsioidea)
  names(Tarsiiformes) <- c("Tarsioidea")
Simiiformes <- list(Platyrrhini, Catarrhini)
  names(Simiiformes) <- c("Platyrrhini", "Catarrhini")

# Making the suborders into lists containing their infraorders
Strepsirrhini <- list(Lorisiformes, Lemuriformes)
  names(Strepsirrhini) <- c("Lorisiformes", "Lemuriformes")
Haplorrhini <- list(Tarsiiformes, Simiiformes)
  names(Haplorrhini) <- c("Tarsiiformes", "Simiiformes")
```

2. Use double bracket notation to extract the New World monkeys (Parvorder Platyrrhini) from this list into a new variable **nwm**. What is the class of this variable? What is the mode of this variable?
```{r}
nwm <- Haplorrhini[[2]][1]
class(nwm) # A: list
mode(nwm) # A: list
```

3. Use `$` notation to extract the tarsiers (i.e., Family Tarsiidae)
```{r}
Haplorrhini$Tarsiiformes
```
